---
title: "lab5"
author: "group12"
date: "11/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(boot)
library(ggplot2)
data <- read.csv2('lottery.csv')
```

# Question 1 **Hypothesis testing**

## Using losse()

Using `loess()`function to get estimate $\hat Y$ , and calculate the statistics

$$
T = \frac{\hat Y(X_b)-\hat Y(X_a)}{X_b-X_a},\ \ \ \ where X_b=argmax_x\hat Y(X),X_a=argmin_x\hat Y(X)
$$

The T-statistic is below. Since it is significantly different from 0, We can not say the lottery is random.

```{r, echo=FALSE}
df <- data.frame(x = data[,4],y = data[,5])
los <- loess(y~x,data = df)
y_hat <- los[['fitted']]
Xb = df$x[which.max(y_hat)]   
Xa = df$x[which.min(y_hat)]
T_ <- (predict(los,Xb)-predict(los,Xa))/(Xb-Xa) 
print(T_)
```

## Distribution of T

Using the bootstrap algorithm(from course pdf files), we have the plot where the red line means 95% significance. This is the hypothesis test where H0: T=0 (lottery is random)

```{r echo=FALSE}
stats <- function(data,vec){
 datatemp<-data[vec,]
  los = loess(y ~ x, data = datatemp)
  res <- predict(los,datatemp)
  Xb = datatemp$x[which.max(res)]   
  Xa = datatemp$x[which.min(res)]
  y_Xb = predict(los,newdata=Xb)
  y_Xa = predict(los,newdata=Xa)
  T_stat = (y_Xb - y_Xa) / (Xb-Xa)
  return(T_stat)
}
set.seed(12345)
myboot = boot(data = df, 
              statistic = stats,
              R = 2000)
print('summary for bootstrap')
myboot

df1 <- data.frame(t=myboot$t)
per95 = sort(myboot$t)[1950]
per95l = sort(myboot$t)[50]
ggplot(data = df1, aes(x = t)) + 
  ggtitle("Histogram of T") + 
  geom_histogram(aes(y=..density..),
                 colour="black",
                 fill="white",
  )+
  geom_vline(aes(xintercept = per95, color = "red"),size=1.5)+
   geom_vline(aes(xintercept = per95l, color = "red"),size=1.5)+
  theme(plot.title = ggplot2::element_text(hjust=0.5))
```

Since x = 0 is out the 95% significance region, we can say that the lottery is not random.

# Question 2 **Bootstrap, jackknife and confifidence intervals**

## 1 Plot and show mean value

This plot can not remind us any familiar distribution.

```{r, echo=FALSE}
rm(list=ls())
data <- read.csv2('prices1.csv')
df <- data.frame(p = data[,1])
ggplot(data = df, aes(x = p)) + 
  ggtitle("Histogram of price") + 
  geom_histogram(aes(y=..density..),
                 colour="black",
                 fill="white",
                 bins=30) + 
  theme(plot.title = ggplot2::element_text(hjust=0.5))
print(paste0('mean value of peice is ', mean(df[,1])))
```

## 2 Compute some statistics

From the course document, we know that

$$
bias-correction = T_1 = 2T(D)-\frac{\sum_{i=1}^B T_i^*}{B}\\
variance\ of\ estimator = \hat {Var[T(.)]} = \frac{\sum_{i=1}^B (T(D^*)-\overline
{T(D^*)})^2}{B-1}
$$

```{r, echo=FALSE}
stat1 <- function(vec,vn){
  return(mean(vec[vn]))
}
B=1000
set.seed(12345)
res1 = boot(data$Price, stat1, R=B)
print('summary of bootstrap')
res1
print('                                                   ')
print(paste0('bias-correction : ',2*res1$t0-mean(res1$t)))
# variance of mean price (output of statistic)
var_boot <- 1/(B-1)*sum((res1$t-mean(res1$t))^2 )
print(paste0('variance : ',var_boot))
print('                                                   ')
# default is a 95% confidence interval
ci <- boot.ci(res1, type = c("perc", "bca", "norm"))
print(ci) 
plot(res1)
```

## 3 About estimate mean

```{r, echo=FALSE}
print(paste0('estimated mean is :',mean(res1$t)))
```

The estimated mean is 1080.853, it located in all confidence intervals.

## 4 jackknife

First, we have the knowledge from the course documents.

$$
\hat{Var[T(.)]} = \frac{1}{n(n-1)}\sum_{i=1}^n ((\rm T_i^*)- \rm J(T))^2,\\
where\ T_i^* = nT(D)-(n-1)T(D_i^*)\\
and \ J(T)= \frac{1}{n}\sum_{i=1}^{n}T_i^*
$$

The variance of mean price is showed below and the comparasion is in the table.

```{r, echo=FALSE}
n = nrow(data)
constant = 1/(n*(n-1))
T_i = sapply(1:n, function(i){
  n * mean(data$Price) - (n-1) * mean(data[-i,1])
})
J_T = (1/n) * sum(T_i)
Var_jac = constant * sum((T_i - J_T)^2)
Var_jac
table = data.frame(boostrap = var_boot, jackknife= Var_jac)
print(table)
```

# Appendix

```{r,eval=FALSE}
library(boot)
library(ggplot2)
data <- read.csv2('lottery.csv')
############################################################################
# Q1
############################################################################
df <- data.frame(x = data[,4],y = data[,5])
los <- loess(y~x,data = df)
y_hat <- los[['fitted']]
Xb = df$x[which.max(y_hat)]   
Xa = df$x[which.min(y_hat)]
T_ <- (predict(los,Xb)-predict(los,Xa))/(Xb-Xa) 
## print(T_)
stats <- function(data,vec){
 datatemp<-data[vec,]
  los = loess(y ~ x, data = datatemp)
  res <- predict(los,datatemp)
  Xb = datatemp$x[which.max(res)]   
  Xa = datatemp$x[which.min(res)]
  y_Xb = predict(los,newdata=Xb)
  y_Xa = predict(los,newdata=Xa)
  T_stat = (y_Xb - y_Xa) / (Xb-Xa)
  return(T_stat)
}
# non-parametric bootstrap
set.seed(12345)
# dt=data1[order(data1$Draft_No),]
myboot = boot(data = df, 
              statistic = stats,
              R = 2000)
## print('summary for bootstrap')
## myboot
# plot distribution
# plot(myboot, index = 1)
df1 <- data.frame(t=myboot$t)
per95 = sort(myboot$t)[1950]
p1 <- ggplot(data = df1, aes(x = t)) + 
  ggtitle("Histogram of T") + 
  geom_histogram(aes(y=..density..),
                 colour="black",
                 fill="white",
                 bins=30) + 
  geom_vline(aes(xintercept = per95, color = "red"),size=1.5)+
  geom_vline(aes(xintercept = per95l, color = "red"),size=1.5)+
  theme(plot.title = ggplot2::element_text(hjust=0.5))
##################################################################################
# Q2 Jackknife
##################################################################################
rm(list=ls())
data <- read.csv2('prices1.csv')
df <- data.frame(p = data[,1])
p2 <- ggplot(data = df, aes(x = p)) + 
  ggtitle("Histogram of price") + 
  geom_histogram(aes(y=..density..),
                 colour="black",
                 fill="white",
                 bins=30) + 
  theme(plot.title = ggplot2::element_text(hjust=0.5))
## print(paste0('mean value of peice is ', mean(df[,1])))
###########################################################################
#q2
###########################################################################
stat1 <- function(vec,vn){
  return(mean(vec[vn]))
}
B=1000
set.seed(12345)
res1 = boot(data$Price, stat1, R=B)
## print('summary of bootstrap')
## res1
## print('                                                   ')
## print(paste0('bias-correction : ',2*res1$t0-mean(res1$t)))
## variance of mean price (output of statistic)
var_boot <- 1/(B-1)*sum((res1$t-mean(res1$t))^2 )
## print(paste0('variance : ',var_boot))
## print('                                                   ')
# default is a 95% confidence interval
ci <- boot.ci(res1, type = c("perc", "bca", "norm"))
## print(ci) 
## plot(res1)
###########################################################################
#q3
###########################################################################
## print(paste0('estimated mean is :',mean(res1$t)))
###########################################################################
#q4
###########################################################################
n = nrow(data)
constant = 1/(n*(n-1))
T_i = sapply(1:n, function(i){
  n * mean(data$Price) - (n-1) * mean(data[-i,1])
})
J_T = (1/n) * sum(T_i)
Var_jac = constant * sum((T_i - J_T)^2)
## Var_jac
table = data.frame(boostrap = var_boot, jackknife= Var_jac)
## print(table)
```
