
---
title: "Solution to computer exam in Bayesian learning"
author: "Bertil Wegmann"
date: "2021-08-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,results="markup")
```

First load all the data into memory by running the R-file given at the exam
```{r}
rm(list=ls())
source("ExamData.R")
set.seed(1)
```

## Problem 1

### 1b
```{r}
y1 <- c(2.32,1.82,2.40,2.08,2.13)
n <- length(y1)
theta <- rgamma(1e5,shape = 2*n+1,rate = 0.5+sum(y1))
y_tilde <- rgamma(1e5, shape = 2, rate = theta)
plot(density(y_tilde),type="l",main="Posterior distribution",xlab="y_tilde",ylab="")
mean(y_tilde < 1.9)
```
The posterior predictive probability is roughly 0.53. The posterior distribution is plotted above.

### 1c
```{r}
nSim <- 1e5
nWeeks <- 30
WeeklyWeights <- matrix(NA,nSim,nWeeks)
for (i in 1:nSim){
  thetas <- rgamma(nWeeks,shape = 2*n+1,rate = 0.5+sum(y1))
  WeeklyWeights[i,] <- t(rgamma(nWeeks, shape = 2, rate = thetas))
}
ExceedingWeeks <- rowSums(WeeklyWeights > 2.4)
mean(ExceedingWeeks)
```
The expected number of weeks is roughly 10.5.

### 1d
```{r}
ExpectedLoss <- function(a, WeeklyWeights){
  EL <- a + mean(rowSums(WeeklyWeights > 0.9*log(a)))
  return(EL)
}
aGrid <- seq(2,10,by = 0.01)
EL <- rep(NA,length(aGrid),1)
count <- 0
for (a in aGrid){
  count <- count + 1
  EL[count] = ExpectedLoss(a, WeeklyWeights)
}
plot(aGrid, EL, type = "l")
aOpt = aGrid[which.min(EL)] # This is the optimal a
points(aOpt,ExpectedLoss(a=aOpt, WeeklyWeights), col = "red",pch=19)
aOpt
```
The optimal build cost (a) is roughly 7.5.

## Problem 2

### 2a
```{r}
mu_0 <- as.vector(rep(0,8))
Omega_0 <- (1/9)*diag(8)
v_0 <- 1
sigma2_0 <- 9
nIter <- 10000
library(mvtnorm)

X <- as.matrix(X)
PostDraws <- BayesLinReg(y, X, mu_0, Omega_0, v_0, sigma2_0, nIter)
Betas <- PostDraws$betaSample
quantile(Betas[,2],probs=c(0.005,0.995))
```
It is 99 % posterior probability that beta_1 is on the interval (-0.33,1.84).

### 2b
```{r}
Mu_draws <- Betas%*%as.vector(c(1,1,1,0.5,0,1,0,1))
Sigma_draws <- sqrt(PostDraws$sigma2Sample)
median(Sigma_draws/Mu_draws)
```
The median of CV is given by roughly 1.83.

### 2c
```{r}
Diff_Exp_Price <- (Betas[,5]+Betas[,7])-(Betas[,6]+Betas[,8])
plot(density(Diff_Exp_Price),type="l",
     main="Posterior distribution for the difference in expected price",
     xlab="",ylab="")
quantile(Diff_Exp_Price,probs=c(0.025,0.975))

plot(density(Betas[,8]),type="l",main="Posterior distribution",xlab="beta_7",ylab="")
```
There is substantial probability mass that the expected price is higher for apartments in the inner city compared to the south side of the city. A 95 % equal tail credible interval for the difference in expected selling price when x1=1 is equal to [1.09,2.41], which further points in this direction.

The posterior distribution for the difference in slopes of x1 between the south side and neither inner city nor south side (beta_7) has substantial probability mass on both sides of 0, so that the effect from x1 on the selling price y is not likely to be different between the two regions.

### 2d
```{r}
Mu_draws <- Betas%*%as.vector(c(1,-0.5,-0.5,0,0,1,0,-0.5))
plot(density(Mu_draws),type="l",main="Posterior distribution of mu",xlab="mu",ylab="")
mean(Mu_draws>0)
```
The posterior distribution of mu is plotted above. The posterior probability is 0.

### 2e
```{r}
x1_grid <- seq(min(X[,2]),max(X[,2]),0.01)
ypred_draws <- matrix(0,length(x1_grid),2)
for (ii in 1:length(x1_grid)){
  CurrMu <- Betas%*%as.vector(c(1,x1_grid[ii],1,0.5,1,0,x1_grid[ii],0))
  ypred_draws[ii,] <- quantile(rnorm(nIter,CurrMu,Sigma_draws),probs=c(0.025,0.975))
}

plot(x1_grid,ypred_draws[,1],"n",
     main="95 % posterior predictive intervals as a function of x1",
     xlab="x1", ylab="",ylim=c(-4,6))
lines(x1_grid,ypred_draws[,1],col="blue")
lines(x1_grid,ypred_draws[,2],col="blue")
```
The posterior predictive intervals as a function of x1 are plotted above.

## Problem 3

### 3d
```{r}
LogPost <- function(theta,n,SumLogx){
  
  logLik <- -n*theta**2 + 2*theta*SumLogx;
  
  return(logLik)
}
theta_grid <- seq(-0.7,1.5,0.001)
PostDens_propto <- exp(LogPost(theta_grid,5,2))
PostDens <- PostDens_propto/(0.001*sum(PostDens_propto))
plot(theta_grid,PostDens,main="Posterior distribution",xlab="theta", ylab="")
```
The posterior distribution is given above.


