---
title: "lib4linerVignette"
author: "Wuhao Wang Xiaolin Jiang"
date: "`2021/09/07 15:35:00"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lib4linerVignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Description

This package offers a RC class for liner regression.This class offers the methods
for computing *fitted_value*,*regression coefficient*,*residuals*,*the residual variance*,
*the variance of the regression coefficients* and *t-values for each coefficient*.

# Import

```{r}
    library(lab4liner)
```

# Methods

The example below is to introduce how these functions work
`linreg()` takes two parameter, object : a formula object ; data :  a data.frame object 

```{r}
lab4lm <- linreg(object = Petal.Length~Species, data = iris)
```

The `print()` prints out the vector containing coefficients.

```{r}
lab4lm$print()
```

The `plot()`  plots such pictures(package ggplot2 is needed here)

```{r}
lab4lm$plot()
```

The `resid()` returns the vector of residuals.

```{r}
head(lab4lm$resid())
```

The `pred()` returns the predicted values.

```{r}
head(lab4lm$pred())
```

The `coef()` returns the coefficients as a named vector.

```{r}
lab4lm$coef()
```

The `summary()`  returns a similar printout as printed for lm objects, but you only need to present
the coefficients with their standard error, t-value and p-value as well as the estimate of ˆσ and the degrees
of freedom in the model.

```{r}
lab4lm$summary()
```