---
title: "lab1computationalsta"
author: "Wuhao Wang"
date: "11/2/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 

q1: Check the result of the snippets. Comment what is going on

In the first test, we get 1/3 - 1/4 != 1/12. That's because computer can not 
store numbers like 1/3 which have infinite digits. So the actual number in the memory
is just very closed to 1/3. As for 1/4,has the finite digits(within 32), computer
can store it without losing accuracy. As a result, the final result of 1/3 - 1/4
is not equal to 1/12.

q2: do you have any problems, suggest improvement？


```{r cars}

x1<- 1/3  
x2<- 1/4
if ( x1-x2==1/12 ) {
print( 'Subtraction is correct' )
} else {
print( 'Subtraction is wrong' )
}

x1<- 1 
x2<- 1/2
if ( x1-x2==1/2 ) {
print( 'Subtraction is correct' )
} else {
print( 'Subtraction is wrong' )
}

```

## Question 2
```{r pressureE}

dev <- function(x)
{
  eta <- 10**-15
  return ((x+eta-x)/eta)
}
print(dev(1))
print(dev(100000))
```
q3 :What values did you obtain? What are the true values? Explain the reasons behind the
discovered differences.

We get 1.110223 at x=1 and 0 at x=100000.

Assume that ETA = 10**-15. When doing this calculating,computer first calculates x+ETA and get an answer(Y),then Y-eta,finally,the division.
In this process,when x==1, x+ETA will lose some part of the mantissa, so Y is not 1+ETA but very closed. So the final result is biased. When
x == 100000, because of the same reason, computer can not store any digits of the ETA part, so here Y == 100000, thus (Y-100000)/ETA = 0

## Question 3 

```{r pressure}
myvar <- function(x)
{
  len <- length(x)
  return((sum(x*x)-sum(x)*sum(x)/len)/(len-1))
}

myvar2 <- function(x)
{
  m <-mean(x)
  len <- length(x)
  return(sum((x-m)*(x-m))/(len-1))
}

test_myvar <-function(x,i)
{
  return(myvar(x[1:i])-var(x[1:i]))
}

test_myvar2 <-function(x,i)
{
  return(myvar2(x[1:i])-var(x[1:i]))
  
}
xq3 <- rnorm(10000,mean=10**8,sd=1)
y<-c()
for(i in 1:10000)
{
  y<-c(y,test_myvar(xq3,i))
}
plot(1:10000,y[1:10000],main = 'original method')

y<-c()
for(i in 1:10000)
{
  y<-c(y,test_myvar2(xq3,i))
}
plot(1:10000,y[1:10000],main = 'promoted method')
```
q3: How well does your function work? Can you explain the behaviour?

It doesn't work well.

In the section sum(x)*sum(x),the integer overflow happens. So the answer will lose
accuracy.

q4: How can you better implement a variance estimator? Find and implement a formula that
will give the same results as var()?

We can use another formula, var = fill the formula here(that one using x-mean(x))


## Question 4
```{r pressureE}
plotQ4 <- function()
{
  max_for_An <- 0 
  max_for_Bn <- 0 
  max_for_Cn <- 0 
  max_for_Bk <- 0 
  max_for_Ck <- 0
  for(i in 1:200)
  {
    if(prod(1:i)==Inf)
    {
      max_for_An <- i
      break
    }
  }
  flagB <- 1
  for(i in 1:200)
  {
    for(j in 0:i)
    {
      if(flagB)
      if(prod((j+1):i)==Inf||prod(1:(i-j))==Inf)
      {
        max_for_Bn <- i
        max_for_Bk <- j
        flagB <- 0 
      }
      if(prod(((j+1):i)/(1:(i-j)))==Inf)
        {max_for_Cn <- i;max_for_Ck <- j;break} 
    }
  }
  res <- c(max_for_An,max_for_Bn,max_for_Cn)
  res2 <- c(0,max_for_Bk,max_for_Ck)
  
  df <-data.frame(method = c('method A','method B','method C'),n = res)
  p1<-ggplot2::ggplot(df, ggplot2::aes(x = method, y = n)) +
  ggplot2::geom_bar(stat = "identity")+
  ggplot2::geom_text(ggplot2::aes(label = n), vjust = 2, colour = "white", position = ggplot2::position_dodge(.9), size = 5)+
  ggplot2::ggtitle('n when overflow occurs')+
    ggplot2::theme(plot.title = ggplot2::element_text(hjust=0.5))
  
  df <-data.frame(method = c('method A','method B','method C'),n = res2)
  p2<-ggplot2::ggplot(df, ggplot2::aes(x = method, y = n)) +
  ggplot2::geom_bar(stat = "identity")+
  ggplot2::geom_text(ggplot2::aes(label = n), vjust = 2, colour = "white", position = ggplot2::position_dodge(.9), size = 5)+
  ggplot2::ggtitle('k when overflow occurs')+
    ggplot2::theme(plot.title = ggplot2::element_text(hjust=0.5))
  print(p1)
  print(p2)
}
plotQ4()
```
Consideration: which computes the product of all the elements of the vector passed to it?

When n and k are comparatively small, all these three methods will compute all the input vector.
However,with n and k become greater and greater,method A will fail to compute first then method B
will also lose its function.Finally, method C will also fail.

q1: Even if overflow and underflow would not occur these expressions will not work correctly
for all values of n and k. Explain what is the problem in A, B and C respectively.

A: when k==0 or k==n, it will compute prod(1:0) and get a zero divisor leading to an Inf result.

B: when n==k, it will compute prod(1:0) and get a zero divisor leading  to an Inf result.

C: when n==k, it will compute prod(1:0) and get a zero divisor leading  to an Inf result.

q2: In mathematical formula one should suspect overflow to occur when parameters, here
n and k, are large. Experiment numerically with the code of A, B and C, for different
values of n and k to see whether overflow occurs. Graphically present the results of your
experiments.

(hits : we assume that n grows from 1 and k grows from 0. In method A, zero divisor overflow will always occur when k==0, so for method A,we only consider n)

From the result we can see that:

In method A, when n = 171, it will overflow.

In method B, when n = 171, k = 0, it will overflow.

In method C, when n = 200, k = 200, it will overflow.

q3: Which of the three expressions have the overflow problem? Explain why.

Method A have most serious overflow problems. Since it will compute prod(1:n) first,then no matter the magnitude of k, it will overflow when n grows.

Method B is better than A, but when K == 0,it will face the same problem as method A.

Method C is the best in these three method,since it compute ((k+1):n)/(1:(n-k)) first,which lower the scale and the risk of overflow.But it inevitably faces overflow problems when n 
and k grow in large scale.

